Discriminating the Influence of Correlated Factors \\from Multivariate Observations: the Back-to-Back Regression

### 2020-05-20 paper/introduction.tex
up

Natural sciences are tasked to find, from a set of hypothetical factors, the
minimal subset that suffices to reliably predict novel observations. This
endeavor is impeded by two major challenges.

First, causal and non-causal factors may be numerous and partially correlated.
% In physics,
% for example, one may be challenged to identify whether fusion is caused by a
% change in temperature or a change in pressure, as these two factors may, at
% first, be difficult to manipulate independently. This issue becomes increasingly
% pronounced as the number of potential factors increases.
In neuroscience, for
example, it can be challenging to identify whether word frequency modulates
brain activity during reading. Indeed, the
frequency of words in natural language covaries with other factors such as their
length (short words are more frequent than long words) and their categories
(determinants are more frequent than adverbs)
\citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words
that controls for all of these factors simultaneously, it is thus common to use
a \emph{forward} "encoding model", i.e. to fit a linear regression to predict observations
(e.g. brain activity) from a minimal combination of competing factors (e.g.
word length, word frequency), and analytically investigate
the estimated contribution of each factor from the model's coefficients
\citep{friston1994statistical,naselaris2011encoding,weichwald2015causal,
king2018encoding,huth2016natural}.

The second challenge to measuring causal influence is that observations can be
multidimensional.
For example, brain activity is often
recorded with hundreds or thousands of simultaneous measurements via functional
Magnetic Resonance Imaging, magneto-encephalography (MEG) or multiple electro-
physiological probes \citep{friston1994statistical,steinmetz2018challenges}.
The relationship between putative causes and observations is thus often
done by training models in a \emph{backward}
fashion: i.e. from observations to putative causal factors. For example, it is
common to fit a support vector machine across multiple
brain voxels or multiple electrodes to detect the
category of a stimulus independently of common noise sources (e.g. head
movements, eye blink)
\citep{norman2006beyond,cichy2014resolving,
kriegeskorte2008representational, king2018encoding}.

Both \emph{forward} and \emph{backward} modeling have competing benefits and drawbacks.
Specifically, forward modeling disentangles the independent contribution of
correlated factors, but does not combine multidimensional observations. By
contrast, backward modeling combines multiple observations, but does not
disentangle factors that are linearly correlated \citep{weichwald2015causal,
hebart2018deconstructing, king2018encoding}. To combine some of the benefits of forward
and backward modeling, several authors have proposed to use cross-decomposition
techniques such as Partial Least Squares (PLS) and Canonical Correlation
Analysis (CCA) \citep{de2019multiway, bilenko2016pyrcca}. CCA and PLS aim to find, from two sets of
data $X$ and $Y$, the components $H$ and $G$ such that $XH$ and $YG$ are maximally
correlated or maximally covarying respectively.

While CCA and PLS can make use
of multidimensional features and observations, they are not designed
for feature discovery. First, these methods are not not directional: observations
and factors can be assigned to either $X$ or $Y$. Second, these method project $X$
and $Y$ onto a reduced but nonetheless multidimensional space. Third, because
CCA and PLS are based on a generalized eigen decomposition, their resulting
coefficients mix the features of $X$ and $Y$ in a way that makes them notoriously difficult to
interpret \citep{lebart1995statistique}.

Here, we introduce the `back-to-back regression' (B2B), which not only combines
the benefits of forward and backward modeling (Section~\ref{sec:algorithm}), but
can also provide robust, interpretable, unidimensional and unbiased coefficients for
each factor.

After detailing B2B and proving its convergence
(Section~\ref{sec:theorem}), we show with synthetic data that it outperforms
state-of-the-art forward, backward and cross-decomposition techniques in
disentangling causal factors (Section~\ref{sec:experiment_synthetic}). Finally,
we apply B2B to large neuroimaging datasets and reveal that distinct but
linearly-correlated word features lead to distinguishable brain representations
(Section~\ref{sec:experiment_real}).

>>>>

Natural sciences are tasked to find, from a set of hypothetical factors, the
minimal subset that suffices to reliably predict novel observations. This
endeavor is impeded by two major challenges. First, causal and non-causal
factors may be numerous and partially correlated.
% In physics,
% for example, one may be challenged to identify whether fusion is caused by a
% change in temperature or a change in pressure, as these two factors may, at
% first, be difficult to manipulate independently. This issue becomes increasingly
% pronounced as the number of potential factors increases.
In neuroscience, for
example, it can be challenging to identify whether word frequency modulates
brain activity during reading. Indeed, the
frequency of words in natural language covaries with other factors such as their
length (short words are more frequent than long words) and their categories
(determinants are more frequent than adverbs)
\citep{kutas2011thirty,pegado2014timing}. Instead of selecting a set of words
that controls for all of these factors simultaneously, it is thus common to use
a \emph{forward} "encoding model", i.e. to fit a linear regression to predict observations
(e.g. brain activity) from a minimal combination of competing factors (e.g.
word length, word frequency), and analytically investigate
the estimated contribution of each factor from the model's coefficients
\citep{friston1994statistical,naselaris2011encoding,weichwald2015causal,
king2018encoding,huth2016natural}.

The second challenge for measuring causal influence is that observations can be
multidimensional.
For example, brain activity is often
recorded with hundreds or thousands of simultaneous measurements via functional
Magnetic Resonance Imaging, magneto-encephalography (MEG) or multiple electro-
physiological probes \citep{friston1994statistical,steinmetz2018challenges}.
The relationship between putative causes and observations is thus often
done by training models in a \emph{backward}
fashion: i.e. from observations to putative causal factors. For example, it is
common to fit a support vector machine across multiple
brain voxels or multiple electrodes to detect the
category of a stimulus \citep{norman2006beyond,cichy2014resolving,
kriegeskorte2008representational, king2018encoding}. Decoding has become
particularly popular in neuroscience, because brain recordings are typically
corrupted by major physiological noise, such as muscle movements, eye blinks,
displacements etc. As these noises sources often lie along specific
components of the multidimensional recordings, the remaining neural signals can
be robustly picked up by multiple decoders \citep{king2018encoding}.

Both \emph{forward} and \emph{backward} modeling have competing benefits and drawbacks.
Specifically, forward modeling disentangles the independent contribution of
correlated factors but does not efficiently combine high-dimensional observations. By
contrast, backward modeling combines multiple observations but does not
disentangle factors that are linearly correlated \citep{weichwald2015causal,
hebart2018deconstructing, king2018encoding}. To combine some of the benefits of forward
and backward modeling, several authors have proposed to use cross-decomposition
techniques such as Partial Least Squares (PLS) and Canonical Correlation
Analysis (CCA) \citep{de2019multiway, bilenko2016Pyrcca}. CCA and PLS aim to find, from two sets of
data $X$ and $Y$, the matrices $H$ and $G$ such that $XH$ and $YG$ are maximally
correlated or maximally covarying respectively.

While CCA and PLS can make use
of multidimensional features and observations, they are not designed
for feature discovery. First, these methods are not not directional: observations
and factors can be assigned to either $X$ or $Y$. Second, these methods project $X$
and $Y$ onto a reduced but nonetheless multidimensional space. Third, because
CCA and PLS are based on a generalized eigen decomposition, their resulting
coefficients mix the features of $X$ and $Y$ in a way that makes them notoriously difficult to
interpret \citep{lebart1995statistique}.

Here, we introduce the `back-to-back regression' (B2B), which not only combines
the benefits of forward and backward modeling (Section~\ref{sec:algorithm}) but
can also provide robust, interpretable, unidimensional and unbiased coefficients for
each factor.

After detailing B2B and proving its convergence
(Section~\ref{sec:theorem}), we show with synthetic data that it outperforms
state-of-the-art forward, backward and cross-decomposition techniques in
disentangling causal factors (Section~\ref{sec:experiment_synthetic}). Finally,
we apply B2B to large MEG and fMRI datasets acquired during a simple reading
task and reveal that B2B efficiently distinguishes the respective effects of
covarying word features (Section~\ref{sec:experiment_real}).

### 2020-05-20 paper/paper.tex
up

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
measure, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it provides a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
least-squares regression and cross-decomposition techniques (e.g. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are partially collinear.
Finally, we apply B2B to magneto-encephalography and functional Magnetic
Resonance Imaging of 102 subjects
recorded during a one hour long reading task to test whether our method appropriately disentangles
the respective contribution of collinear factors such as word length, word
frequency respectively known to cause early visual and late associative cortical
responses. The results
show that these factors are better disentangled with B2B than with
other standard techniques.
\end{abstract}

\begin{keyword}
  Cross-Decomposition \sep
  Feature Discovery \sep
  MEG \sep
  fMRI \sep
  Decoding \sep
  Encoding \sep
  Reading \sep
\end{keyword}

@@

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{m \times
d_y}$ (the dependent variables), generated from a set of putative causes $X \in \mathbb{R}^{m \times
d_x}$ (the independent variables), via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$.
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $S
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XS$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i S + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific additive noise drawn from a
centered distribution.
%
While $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $m$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Figure~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix $S$, i.e. to identify the factors
that most reliably account for the multivariate observations.

@@

Performing two regressions over the same data sample can result in overfitting,
as spurious correlations in the data absorbed by the first regression will be
leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two splits $(X_1, Y_1)$
and $(X_2, Y_2)$.
%
The first regression is performed using $(X_1, Y_1)$, and the second
regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, the two
successive regressions are
repeated over many random splits, and the final estimate $\hat S$ of the causal
influence matrix is the average over the estimates associated to each split
\citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we use an efficient
leave-one-out cross-validation scheme as detailed in \citep{rifkin2007notes}
% http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf,
as follows:
%

\begin{equation}
\hat{Y} = (\Sigma_X G Y - \text{diag}(\Sigma_X G) Y) \;/\; \text{diag}(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

@@

Note that $\S hat$ is unbiased, in the sense that it is centered around zero when there is no
effect, only if the second regression $H$ is not regularized. Second-level
statistics testing whether $\S hat$ is superior to 0 are thus only valid if $H$
is not regularized.

@@

We perform two sets of experiments to evaluate B2B: one on controlled synthetic
data, and a second one on a real, large-scale magneto-encephalography (MEG)
dataset.
%
We use scikit-learn's PLS and RidgeCV \citep{sklearn} as well as Pyrcca's
regularized canonical component analysis (RegCCA, \citep{bilenko2016pyrcca})
objects to compare B2B against the standard baselines.

\subsection{Synthetic data}
\label{sec:experiment_synthetic}
\input{synthetic.tex}

\subsection{functional Magnetic Resonance Imaging}
\label{sec:experiment_fmri}
\input{fmri_exps.tex}

\subsection{Magneto-encephalography}
\label{sec:experiment_real}
\input{meg_exps.tex}

@@

Here, we introduce B2B, a method to disentangle the causal contribution of
collinear factors from multidimensional observations. After proving the validity
of B2B, we show that it compares favorably against baseline models, such as
Backward, Forward, CCA and PLS models both on a wide spectrum of synthetic data,
and on two large neuroimaging datasets.

In addition, B2B can be very fast to compute as long as both $H$ and $G$ are based
on l2 optimization with l2 regularization \ref{fig:duration}. Furthermore, B2B
need not be limited to these standard optimizations: the $H$ and $G$ operators
can in principle be found with e.g. support vector or random forest regression.
We reserve the study of these computational variants to future work.

Like Forward and cross-decomposition models, B2B is limited by the correlations
between factors. At the extreme, if two factors are identical, and thus fully
correlated, none of these statistical method can disentangle the causal
contribution of each of them separately. In practice, this implies that, like
general linear models, B2B will best work with high signal-to-noise ratio and/or
orthogonal factors, and will see its sensitivity diminish when the
signal-to-noise ratio of collinear factors drops.

In the present neuroimaging context, B2B follows a long series of statistical
methods designed to characterize brain representations - i.e. to identify what
sensory feature causes specific brain responses \citep{naselaris2011encoding}.
In this regard, CCA and PLS have been used in electrophysiology and neuroimaging
for tracking representations (e.g. \citep{musch2020transformation}) as well as
for denoising and subject alignment \citep{cca_hotelling, de2019multiway}. While
CCA and PLS relates to B2B, these methods diverge in several important ways.
First, they have different objectives: CCA aims to find the potentially numerous and poorly
interpretable components where $X$ and $Y$ are maximally correlated, whereas B2B
aims to recover the causal factors from X to Y. Second, B2B is not symmetric
between $X$ and $Y$: it aims to identify specific causal features by first
optimizing over the decoders $G$ and then over $H$. By contrast, CCA and PLS are
symmetric between $X$ and $Y$, and aims to find $G$ and $H$ such that they
project $X$ and $Y$ on maximally correlated dimensions. Third, CCA is based an
eigen decomposition of $XH$ and $YG$ - the corresponding canonical components
are thus mixing the $X$ features in way that limit interpretability and
potentially dilute the impact of each feature onto multiple components. In
contrast B2B assesses each feature $X^j$ on a single $Y$ component specifically
selected to maximize signal-to-noise ratio of that feature $j$. Fourth CCA does
not separately optimize two distinct regularization parameters for $G$ and $H$,
whereas B2B does. Finally, CCA does not use different data splits to
estimate $G$ and $H$. Together, these differences may explain why B2B reliably
outperform CCA on estimating causal influences (Figs. \ref{fig:percondition} and
\ref{fig:auc_plots}).

One popular method to investigate multidimensional patterns of brain activity is
Representational Similarity Analysis (RSA)
\citep{kriegeskorte2008representational}. RSA quantifies the
similarity of brain responses associated with specific categorical conditions
(e.g. distinct images), by (1) fitting one-against-all classifiers on each
category and (2) testing whether these classifiers discriminates all other
categories. The resulting categories $\times$ categories confusion matrix is
then analyzed, generally in an unsupervised manner, to reveal the categories
that present similar brain activity patterns. B2B
subsumes RSA in that (1) it uses regressions instead of one-hot
classifications and (2) it is fully supervised.
Consequently, and unlike RSA, B2B (1) provides interpretable coefficients and
(2) can generalize to new items and new contexts. In practice, these elements
allow B2B to apply to event-related paradigms and latent variable analyses,
whereas RSA can only be applied when the same one-hot-encoded condition is
repeated multiple times.

>>>>

\begin{abstract}
Identifying causes solely from observations can be particularly challenging when
i) potential factors are difficult to manipulate independently and ii)
observations are multi-dimensional. To address this issue, we
introduce ``Back-to-Back'' regression (B2B), a linear method designed to efficiently
estimate, from a set of correlated factors, those that most plausibly account for
multidimensional observations. First, we prove the consistency of B2B, its links to
other linear approaches, and show how it can provide a robust, unbiased and
interpretable scalar estimate for each factor.
Second, we use a variety of simulated data to show that B2B outperforms
forward modeling ("encoding"), backward modeling ("decoding") as well as
cross-decomposition modeling (i.e.. canonical
correlation analysis and partial least squares) on causal identification when
the factors and the observations are not orthogonal.
Finally, we apply B2B to a hundred magneto-encephalography recordings and to a hundred functional Magnetic
Resonance Imaging recordings acquired during a one hour reading task. B2B successfully disentangles
the respective contribution of collinear factors such as word length, word
frequency in the early visual and late associative cortical
responses respectively. These factors are better disentangled with B2B than with
other standard techniques. We discuss how the speed and generality of B2B
sets the foundations of an omnibus test for disentanglement of covarying
factors in multidimensional observations.
\end{abstract}

\begin{keyword}
  Feature Discovery \sep
  MEG \sep
  fMRI \sep
  Decoding \sep
  Encoding \sep
  Cross-Decomposition \sep
  Reading \sep
\end{keyword}

@@

We consider the measurement of multivariate signal $Y \in \mathbb{R}^{m \times
d_y}$ (the dependent variables, e.g. the neural responses),
generated from a set of putative causes $X \in \mathbb{R}^{m \times
d_x}$ (the independent variables, e.g. the features of a stimulus),
via some unknown linear apparatus $F \in \mathbb{R}^{d_x \times d_y}$ (e.g. the
projection of neural activity onto MEG channels or fMRI voxels).
%
Not all the variables in $X$ exert a causal influence on $Y$.
%
By considering a square binary diagonal matrix of \emph{causal influences} $S
\in \mathbb{D}^{d_x \times d_x}$, we denote by $XS$ the causal factors of $Y$.
%
In summary, the problem can be formalized as:
%
\begin{equation}
    y_i = (x_i S + n_i) F
    \label{eq:model}
\end{equation}
%
where $i$ is a given sample, and $n_i$ is a sample-specific additive noise drawn from a
centered distribution.
%
While $X$ and $N$ are independent, we allow each of
them to have any form of covariance.
% what is a general covariance matrix?
%
In practice, we observe $m$ samples $(X, Y)$ from the model.
%
This problem space, along with the sizes of all variables involved, is
illustrated in Fig.~\ref{fig:b2b}.
%
Given the model in Equation~\cref{eq:model}, the goal of Back-to-Back
Regression (B2B) is to estimate the matrix $S$, i.e. to identify the factors
that reliably account for the multivariate observations.

@@

Performing two regressions over the same data sample can result in overfitting,
as spurious correlations in the data absorbed by the first regression will be
leveraged by the second one.
%
To avoid this issue, we split our sample $(X, Y)$ into two disjoint sets $(X_1, Y_1)$
and $(X_2, Y_2)$.
%
The first regression is performed using $(X_1, Y_1)$, and the second
regression is performed using $(X_2, Y_2)$.
%
To compensate for the reduction in sample size caused by the split, the two
successive regressions are
repeated over many random splits, and the final estimate $\hat S$ of the causal
influence matrix is the average over the estimates associated with each split
\citep{breiman1996bagging}.
%
To accelerate this ensembling procedure, we use an efficient
leave-one-out (LOO) cross-validation scheme as detailed in \citep{rifkin2007notes}
% http://cbcl.mit.edu/publications/ps/MIT-CSAIL-TR-2007-025.pdf,
as follows:
%

\begin{equation}
\hat{Y}_{LOO} = (\Sigma_X G Y - \text{diag}(\Sigma_X G) Y) \;/\; \text{diag}(I - \Sigma_X G) \qquad \text{(element-wise division)}
\end{equation}

@@

Note that $hat \S $ is unbiased, in the sense that it is centered around zero when there is no
effect, only if the second regression $H$ is not regularized. Second-level
statistics testing whether $hat \S$ is superior to 0 are thus only valid if $H$
is not regularized.

@@

We perform three sets of experiments to evaluate B2B: one on controlled synthetic
data, a second one on a real, large-scale functional Magnetic Resonance Imaging
(fMRI) dataset and a third one on a real, large-scale magneto-encephalography (MEG)
dataset.
%
We use scikit-learn's PLS and RidgeCV \citep{sklearn} as well as Pyrcca's
regularized canonical component analysis (RegCCA, \citep{bilenko2016Pyrcca})
objects to compare B2B against the standard baselines, with common
hyper-parameter optimizations.

\subsection{Synthetic Experiment}
\label{sec:experiment_synthetic}
\input{synthetic.tex}

\subsection{functional Magnetic Resonance Imaging Experiment}
\label{sec:experiment_fmri}
\input{fmri_exps.tex}

\subsection{Magneto-encephalograph Experiment}
\label{sec:experiment_real}
\input{meg_exps.tex}

@@

Here, we introduce B2B, a method to disentangle the causal contribution of
collinear factors from multidimensional observations. After proving the validity
of B2B, we show that it generally compares favorably against baseline models
both on a wide spectrum of synthetic data and on two large neuroimaging datasets.

In addition, B2B can be very fast to compute as long as both $H$ and $G$ are based
on l2 optimization with l2 regularization (Supplementary
Fig.~\ref{fig:duration}), as is done in the above experiments. However, B2B
does not need not be limited to l2 optimization and regularization: the $H$ and $G$ operators
can in principle be found with e.g. support vector or random forest regression.
We reserve the study of these computational variants for future work.

Like forward and cross-decomposition models, B2B is limited by the correlations
between factors. At the extreme, if two factors are identical, and thus fully
correlated, no statistical method can disentangle their relative causal
contribution, and intervention is thus mandatory. In practice, this implies that, like
general linear models, B2B will best work with high signal-to-noise ratio and/or
orthogonal factors, and will see its sensitivity diminish when the
signal-to-noise ratio of collinear factors drops.

In the present neuroimaging context, B2B follows a long series of statistical
methods designed to characterize brain representations - i.e. to identify what
sensory feature causes specific brain responses \citep{naselaris2011encoding}.
In this regard, CCA and PLS have been used in electrophysiology and neuroimaging
to track representations (e.g. \citep{musch2020transformation}) as well as
to denoise recordings as well as to align subjects \citep{cca_hotelling, de2019multiway}. While
CCA and PLS relates to B2B, these methods diverge in several important ways.
First, they have different objectives: CCA aims to find the potentially numerous and poorly
interpretable components where $X$ and $Y$ are maximally correlated, whereas B2B
aims to recover the causal factors from X to Y. Second, B2B is not symmetric
between $X$ and $Y$: it aims to identify specific causal features by first
optimizing over the decoders $G$ and then over $H$. By contrast, CCA and PLS are
symmetric between $X$ and $Y$, and aims to find $G$ and $H$ jointly such that they
project $X$ and $Y$ on maximally correlated dimensions. Third, CCA is based an
eigen decomposition of $XH$ and $YG$ - the corresponding canonical components
are thus mixing the $X$ features in way that limit interpretability and
potentially dilute the impact of each feature onto multiple components. In
contrast B2B assesses each feature $X_i$ on a single $Y$ component specifically
selected to maximize signal-to-noise ratio of that feature $i$. Fourth CCA does
not separately optimize two distinct regularization parameters for $G$ and $H$,
whereas B2B does. Finally, CCA does not use different data splits to
estimate $G$ and $H$. Together, these differences may explain why B2B generally
outperform CCA on estimating causal influences (Figs.~\ref{fig:percondition} and
\ref{fig:auc_plots}).

One popular method to investigate multidimensional patterns of brain activity is
Representational Similarity Analysis (RSA)
\citep{kriegeskorte2008representational}. RSA quantifies the
similarity of brain responses associated with specific categorical conditions
(e.g. distinct images), by (1) fitting one-against-all classifiers on each
category and (2) testing whether these classifiers discriminates all other
categories. The resulting categories $\times$ categories confusion matrix is
then analyzed, generally in an unsupervised manner, to reveal the categories
that present similar brain activity patterns. B2B
subsumes RSA in that (1) it can use regressions instead of one-hot
classifications and (2) it is fully supervised.
Consequently, and unlike RSA, B2B (1) provides interpretable coefficients and
(2) can generalize to new items and new contexts. In practice, these elements
allow B2B to apply to event-related paradigms and latent variable analyses,
whereas RSA can only be applied when the same one-hot-encoded condition is
repeated multiple times.

### 2020-05-20 paper/synthetic.tex
up

All baseline methods were based on the implementations of scikit-learn \citep{sklearn} and pyrcca
\citep{bilenko2016pyrcca}. For pedagogical purposes, we briefly summarize them below.

@@

We employ five-fold cross-validation to select the optimal number of components
for CCA and PLS. Regressions were $\ell2$-regularized with a $\lambda$ regularization
parameters fitted with the efficient leave-one-out procedure implemented in
scikit-learn RidgeCV \citep{sklearn}.

@@

%
B2B leads to \emph{scalar} coefficients for non-causal
features. The diagonal of this matrix, $\hat S$, can thus be directly used
as a causal contribution estimate. Note that this estimate is unbiased (i.e.
zeros-centered) only if the second regression $H$ is not regularized.

In contrast, the Forward, Backward, CCA and PLS models lead to a
loading \emph{vector} for each feature: each vector indicates how to a feature
relates to each dimension of the observations (for Forward or Backward
models) or each canonical projection of the observations.
To estimate causal contribution, we must thus transform such vector into a
scalar, by e.g. taking the sum of square coefficients:
% \begin{equation}
  $\hat S_i = \sum_j {H^j_i}^2 $
% \end{equation}
Note that in such B2B cases, the estimates are  thus necessarily positive and thus biased.

@@

@@

\subsubsection{Evaluating Causal Discovery with held-out prediction reliability}

In most real-world cases, $S$ is not known. Consequently, the above AUC evaluation can not be estimated.
To address this issue, we assess the ability of each model to reliably predict
independent and identically distributed data from $Y$, given all of the $X$
features versus all-but-one feature $X_{-i}$ (i.e. 'knock-out X'). This
procedure results in two correlation metrics $R_{full}$ and $R_{knockout}$ for
each feature (for the B2B and Backward models) for each dimension of $Y$ (for
the Forward model) or each canonical dimension of $Y$ (for CCA and PLS). The
difference $\Delta R_i = R_{full}-R_{knockout}$ indicates how much each
$X_i$ improves the prediction of (a pre-defined dimension or projection of) $Y$.

We show in Appendix~\ref{appendix:feature_importance} pseudo-code to assess
feature importance for each model. For the Backward Model, feature importance
cannot be assessed as the $X$ collinearity is never
taken into account.

\subsubsection{Results of the synthetic experiment}

Figures~\ref{fig:percondition} (bottom) and \ref{fig:auc_plots} (right, in
Appendix) show the results of our synthetic experiment. Note that the figure
does not display each feature separately, as they are
randomly generated. $\Delta R$ represents the average of $\Delta R_i$ across
features $i$.
Overall both the AUC and the held-out prediction reliability evaluations
show that B2B compares favorably to the baseline models.

>>>>

All baseline methods were based on the implementations of scikit-learn \citep{sklearn} and Pyrcca
\citep{bilenko2016Pyrcca}. For pedagogical purposes, we briefly summarize them below.

@@

We employ five-fold nested cross-validation to select the optimal number of components
for CCA and PLS. Regressions were $\ell2$-regularized with a $\lambda$ regularization
parameters fitted with the efficient leave-one-out procedure implemented in
scikit-learn RidgeCV \citep{sklearn}.

@@

%
B2B leads to \emph{scalar} coefficients for non-causal
features. The diagonal of this matrix, $\hat S \in \mathbb{R}^{d_x}$, can thus be directly used
as a causal contribution estimate. Note that this estimate is unbiased (i.e.
zeros-centered) only if the second regression $H$ is not regularized.

In contrast, the loading coefficients of the Forward ($H_i \in \mathbb{R}^{d_y}$),
Backward ($G^i \in \mathbb{R}^{d_y}$), CCA and PLS models ($H_i \in \mathbb{R}^{d_z}$) lead to a
loading \emph{vector} for each feature $i$.
To estimate causal contribution, we must thus transform such vectors into
scalars, by e.g. taking the sum of square coefficients:
% \begin{equation}
  $\hat S_i = \sum_j {H^j_i}^2 $
% \end{equation}
Note that in such B2B cases, the estimates are thus positive and would thus bias
a second-level statistical analysis against 0.

@@

Figures~\ref{fig:percondition} (top) and~\ref{fig:auc_plots} (top) show the
results of this AUC evaluation. Note that the figure does not display each
feature separately, as they are randomly generated. The results show that B2B
compares favorably to other methods on these synthetic data.

@@

\subsubsection{Evaluating Causal Discovery through the reliability of held-out prediction}

In most real-world cases, $S$ is not known. Consequently, the above AUC
evaluation cannot be estimated.
To address this issue, we assess the ability of each model to reliably predict
independent and identically distributed data from $Y$, given all of the $X$
features versus all-but-one feature $X_{-i}$ (i.e. 'knock-out X'). This
procedure results in two correlation metrics $R_{full}$ and $R_{knockout}$ for
each feature (for the B2B and Backward models) for each dimension of $Y$ (for
the Forward model) or each canonical dimension of $Y$ (for CCA and PLS). The
difference $\Delta R_i = R_{full}-R_{knockout}$ indicates how much each
$X_i$ improves the prediction of
a) the target dimension (i.e. $G^iY \in \mathbb{R}$ for B2B,
b) the average across all of the dimensions $j$ of $Y$
($\frac{1}{d_y}\sum_{j}^{d_y}\Delta R_i^{j}$) for the
Forward model or
c) the average across the canonical dimensions $j$ of $Y$
($\frac{1}{d_z}\sum_{j}^{d_z}\Delta R_i^{j}$) for CCA and PLS.
We show in Appendix~\ref{appendix:feature_importance} pseudo-code to assess
feature importance for each model. For the Backward Model, feature importance
cannot be assessed as the $X$ collinearity is never
taken into account.

Figures~\ref{fig:percondition} (bottom) and~\ref{fig:auc_plots} (right, in
Appendix) show the results of this evaluation on held-out data.
Overall both the AUC and the held-out prediction reliability evaluations
show that B2B compares favorably to the baseline models.
